---
title: "kendallkernel - classify gene expression data"
author: "Yunlong Jiao"
date: "26 February 2016"
output: html_document
---

```{r setup, eval = TRUE, message = FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, fig.width = 8, fig.height = 8, dev = "pdf", fig.keep = "high", fig.path = "figure/", cache.path = "cache/")
set.seed(35875954)

# utiles
library(kernlab) # for kernel svm
library(pcaPP) # for fast kendall tau
library(caret) # for data split
library(parallel) # for parallel cross-validation
source("func.R")
dyn.load("src/tsp.so") # for tsp-related C codes
dyn.load("src/utiles.so") # for other C codes
```

## Datasets

All datasets are taken from publicly available sources. Briefly, for datasets with two independent parts (marked `indepval`), predictions are reported on the test set whilst 5-fold cv training is done on training set; for datasets with only one part (marked `crossval`), predictions are reported by 10 times of 5-fold cv whilst (nested) 5-fold cv is done on each training fold for parameter tuning. See [the paper](https://hal.archives-ouvertes.fr/hal-01279273) for detailed summary of the datasets.

```{r data, cache = TRUE}
# datasets
indepvallist <- c('bc', 'lcbwh', 'pc') # two sets are available
crossvallist <- c('ct', 'ocpbsii', 'cns', 'pcout', 'transbig', 'lung_annarbor_outcome', 'Wang_Breastcancer') # only one set is available so that double-nested cv is necessary
prefixlist <- c(indepvallist, crossvallist)

# read in 10 datasets from folder data/
fnames <- list.files(path = "data/")
for (fname in fnames) load(paste0("data/", fname))

# dataset alias as they appear in the paper
cbind(dataset = (namebefore <- c("bc", "transbig", "Wang_Breastcancer", "ct", "lung_annarbor_outcome", 
                                 "lcbwh", "cns", "ocpbsii", "pc", "pcout")), 
      alias = (nameafter <- c("BC1", "BC2", "BC3", "CT", "LA1", "LC2", "MB", "OC", "PC1", "PC2")))
```

## Model performance comparison

Models come from 3 categories (presented in different ways for ease of coding implementation or for ease of scoring and plotting):

1. A baseline model with no tuning parameter that is all-pairs-majority-votes (or n(n-1)/2-TSP)
2. Models involving only tuning C that are SVM with linear, Gaussian RBF, (2nd-order homogeneous) polynomial and Kendall kernel where KFD are penetrated in SVM codes as simple reference kernel machines
3. Models involving tuning C and k that are SVM with top-k pairs of features with aforementioned kernels

```{r param, cache = TRUE}
# set list of C parameters for SVM-based models
Cpara_list <- c(0.01, 0.1, 1, 10, 100, 1000)
names(Cpara_list) <- paste('C',1:length(Cpara_list),sep='')

# set list of #genepairs for corresponding models
max_nodes = 5000; npairs_out = 30;
npairs_list <- floor(exp(seq(0,1,length.out=npairs_out)*log(max_nodes)))
evenidx <- npairs_list %% 2 == 0
npairs_list[evenidx] <- npairs_list[evenidx] - 1 # keep odd numbers only
npairs_list <- unique(npairs_list)
names(npairs_list) <- paste('k',1:length(npairs_list),sep='')

# categorize models for ease of training
modelsNOpara <- c("APMV")
modelsConly <- c("SVMlinearALL", "SVMkdtALL", "SVMpolynomialALL", "SVMrbf") # plus KFD coded within each
modelsTSPrelated <- c("TSP", "kTSP", "SVMlinearTOP", "SVMkdtTOP", "SVMpolynomialTOP")
# OR reorganise for ease of plotting
models0 <- c("TSP", "APMV")
modelsConly <- c("SVMlinearALL", "SVMkdtALL", "SVMpolynomialALL", "SVMrbf") # same as before!
modelsKonly <- c("kTSP")
modelsCandK <- c("SVMlinearTOP", "SVMkdtTOP", "SVMpolynomialTOP")
modelslist <- c(models0, modelsConly, modelsKonly, modelsCandK)
# OR reorganise for feature selection plot varying K
modelsVary <- c("SVMlinearTOP", "SVMkdtTOP", "SVMpolynomialTOP", "kTSP")
modelsStatic <- c("SVMlinearALL", "SVMkdtALL", "SVMpolynomialALL", "TSP", "APMV")
```

```{r indepval, cache = TRUE, message = FALSE}
# indepval datasets
res_indepval <- mclapply(indepvallist, function(prefixname){
  xtr <- get(prefixname)$xtrain; ytr <- get(prefixname)$ytrain
  xtst <- get(prefixname)$xtest; ytst <- get(prefixname)$ytest
  if(is.null(xtst) || is.null(ytst)) stop(paste('dataset error',prefixname,sep=':'))
  
  res <- perfClassification(NULL, prefixname, xtr, ytr, xtst, ytst,
                            Cpara_list, npairs_list, modelsConly, modelsTSPrelated, modelsNOpara,
                            nfolds = 5, nrepeats = 1, seed = 206)
  return(res)
}, mc.cores = 8)
names(res_indepval) <- indepvallist
```

```{r crossval, cache = TRUE, message = FALSE}
# crossval datasets
res_crossval <- mclapply(crossvallist, function(prefixname){
  xtr <- get(prefixname)$xtrain; ytr <- get(prefixname)$ytrain
  xtst <- get(prefixname)$xtest; ytst <- get(prefixname)$ytest
  if(!is.null(xtst) || !is.null(ytst)) stop(paste('dataset error',prefixname,sep=':'))
  
  set.seed(1226)
  outterFoldIndices <- createMultiFolds(1:nrow(xtr), k=5, times=10)
  sig <- sigest(xtr,scaled=F)['50%']
  
  res <- lapply(outterFoldIndices, function(outterFold){
    return(perfClassification(NULL, prefixname, xtr[outterFold,,drop=F], ytr[outterFold], xtr[-outterFold,,drop=F], ytr[-outterFold], 
                              Cpara_list, npairs_list, modelsConly, modelsTSPrelated, modelsNOpara, 
                              nfolds = 5, nrepeats = 1, seed = 206, sigma=sig))
  })
  return(res)
}, mc.cores = 8)
names(res_crossval) <- crossvallist
```

We report classification accuracy across different datasets and different models.

```{r perf_table}
modelsKFD <- sub("SVM", "KFD", modelsConly)
table_acc <- matrix(-100, 
                    nrow = length(prefixlist), ncol = length(c(modelslist,modelsKFD)),
                    dimnames = list(prefixlist, c(modelslist,modelsKFD)))

for (prefixname in prefixlist) {
  for (modelname in modelslist) {
    if (prefixname %in% indepvallist) {
      res <- res_indepval[[prefixname]]
      idx <- which.max(res[[modelname]]$cvacc)
      s <- res[[modelname]]$acc[idx]
      table_acc[prefixname,modelname] <- round(100*s,2)
      if (modelname %in% modelsConly) { # add KFD penetrated within
        s_kfd <- res[[modelname]]$acc_kfd
        table_acc[prefixname,sub("SVM", "KFD", modelname)] <- round(100*s_kfd,2)
      }
    } else if (prefixname %in% crossvallist) {
      s <- mean(sapply(res_crossval[[prefixname]], function(res){
        idx <- which.max(res[[modelname]]$cvacc)
        return(res[[modelname]]$acc[idx])
      }))
      table_acc[prefixname,modelname] <- round(100*s,2)
      if (modelname %in% modelsConly) { # add KFD penetrated within
        s_kfd <- mean(sapply(res_crossval[[prefixname]], function(res){
          return(res[[modelname]]$acc_kfd)
        }))
        table_acc[prefixname,sub("SVM", "KFD", modelname)] <- round(100*s_kfd,2)
      }
    } else {
      stop("Please add ", prefixname, " in either indepvallist or crossvallist")
    }
  }
}
rownames(table_acc) <- nameafter[match(rownames(table_acc), namebefore)] # re-name 
table_acc <- rbind(AVERAGE = round(colMeans(table_acc), 2), table_acc) # add AVERAGE scores over all datasets
table_acc <- table_acc[order(rownames(table_acc)),order(table_acc["AVERAGE",],decreasing = TRUE)] # re-order
# show score table
t(table_acc)
# show boxplot
par(mar = c(10, 5, 1, 1) + 0.1, font.lab = 2, font.axis = 2, font.main = 2, cex.axis = 1.5, cex.lab = 1.5, cex.sub = 1.5)
boxplot(table_acc[-1,]/100, las = 2, ylab = 'acc', col='royalblue2')








# save up for now...
save.image("res.RData")
```

## session info

```{r session_info}
devtools::session_info()
```
